---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

The table categorizes incedences of diffrent types of crime by month. Some of the categories include robbery, theft and homicide. I think the categories that would be most important from a reporting standpoint are any categories that that have had a signifigant upward or downward trend over the months. Some of them like theft from auto, robbery vehicle and auto stolen are similar sounding categories that may have some overlap making it hard for readers to differenciate. 

### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime %>% 
   mutate(
  `total_incidents` =  Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec,
  `average_incidents` = `total_incidents` / 12
   )
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

Accident, auto stolen and theft from auto have the highest averages. Homicide, B & E other and sex offenses have the lowest averages. I would say that this data shows that PG county on average has higher rates of property crime and lower rates of violent crime on average. I would want to know what precentage of these crimes happen concurrently or have the same perpetrator. 


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
`summer_crime_averages` <- pg_crime %>% 
  mutate(
   `june_change` = ( Jun - May) / May,
   `july_change` = (Jul - Jun) / Jun,
   `august_change` = (Aug -Jul) / Jul,
  )
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

Looking at the summer changes I notice that most of the crimes are neither consistantly rising or falling during summer months. Only two categories theft and accident with impound fell consistantly over the three months. I would want to take an average of the rates of change for the summer months to see which way they lean but I wouldnt say that this claim is nessisarily true in PG county based off this data alone.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Auto related crimes had the higest incidence rate of crimes in Prince George's County in 2024. 

Crimes involving accidents, stolen vehicles and theft from vehicles were categories with incidence rates of over 350 per month last year, according to Prince Georges County crime data. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
