---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv")
```

### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

The table displays the name of each employee, their title, position type, and their salary. What stands out to me is that there are some employees making no salary at all that may cause problems with anaylasis. 

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |> 
  summarize(mean_salary = mean(salary),
            median_salary = median(salary)
)
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |> 
  summarize(mean_salary = mean(salary),
            median_salary = median(salary),
            min_salary = min(salary),
            max_salary = max(salary)
)
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The mean and median are relatively similar which shows that theyre are not too many huge outliers. I belive that there being a few 0 salaries which is the minimum has pulled the average salary down, but it is balenced by there also being some salaries like the maximum that are more than double the medium or mean. I would say the data is spred pretty evenly.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

I would want to know more about the job duties of who is making the most and least money and then take a look at the efficacy of their department and see if there are any mismatches. I would also want to know for the higher salaried people if any connections may have helped them obtain their job and what the person who held the position before them made. 
