---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities %>% mutate(
  violent_rate_2019 = (pop2019 / violent2019 * 1000),
  property_rate_2019 = (pop2019 / property2019 * 1000),
  violent_rate_2020 = (pop2020 / violent2020 * 1000), 
  property_rate_2020 = (pop2020 / property2020 * 1000),
)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

In 2019 Baltimore had the higest violent rate and Ocean City had the highest property rate. In 2020 Ocean City had the higest violent and property rate. I think the pattern of Ocean City being in the top, even in 2019 Ocean City still came second for violent crime rate may show that the tansient nature of the vacation community that hosts large numbers of partying young adults thoughout the summer may be making it prime for higher rates of violent and property crimes. 



### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates %>% mutate(
  violent_change = (violent_rate_2020 - violent_rate_2019) / violent_rate_2019,
  property_change = (property_rate_2020 - property_rate_2019) / property_rate_2019,
)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

What I found interesting was that for most cities both violent and property crime decreased in 2020. In many cities decreases were by more than 50%. Im sure this mostly due to covid lockdowns which would be the story here in my opinion I did find it interesting that in in some areas, including baltimore city crime still rose dispite the downward trend. 

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Ocean City led state property and violent crime rates in 2020 despite a fall in crime amid COVID-19 lockdowns.

The city saw a 45% decrease in violent crime and a 15% decrease in property crime rates during 2020, but still had the highest rates in both categories statewide.
### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile? Do the results change your best story idea from the previous assignment? Why or why not?

```{r}
md_cities_percentiles <- md_cities_rates %>% mutate(
  precentile = ntile( violent_rate_2020, 100))
```

20 cities are above the 90th precentile, this would not change my story idea, because by city is still in the 99th precentile for the highest violent rate.  
