---
title: "Multiple Regression Analysis: Maryland Voter Turnout"
name: "Chelsea Collier"
---

# Introduction

For my final project, I am analyzing a dataset that includes all contestants from Seasons 1 through 43 of the reality show Survivor. The dataset covers from Season 1 in 2000 to Season 43 in 2022, and includes demographic details, gameplay outcomes, and personal background information for each contestant.

The original dataset contained over 30 columns, but I’ve cleaned and narrowed it down to the best ones for my research questions. These include:
	•	profession: the contestant’s occupation, which I categorized into broader employment types such as Healthcare, Education, Business, etc.
	•	state: used to assign contestants to U.S. regions (East Coast, Midwest, South, West)
	•	region: derived from state data, this field is key to testing my hypothesis about East Coast performance
	•	finish and normalized_finish: these represent a contestant’s placement in their season, which is the main outcome variable
	•	num_season: the season number, which helps track changes across time
	•	birthdate, age, and gender: optional control variables for further analysis

#Research Questions

My main hypotheses are:
	1.	Contestants in service-related professions (e.g., education, healthcare, hospitality) place higher on average in Survivor
	2.	Contestants from the East Coast perform better in terms of final placement compared to those from other regions
	3.	Age Hypothesis: Contestants aged 25 to 35 are more likely to finish higher in the game compared to other age groups

#What’s Missing
	•	Income or education levels of contestants aren’t included, which might add context to occupation-based patterns
	•	A few state entries are missing or ambiguous, which affects accurate regional classification.
	•	I may also need to explore whether normalized vs. raw finish is the better metric for comparison across seasons with different numbers of contestants

#Questions
	•	Should I analyze raw finish (finish) or scaled placement (normalized_finish) to compare contestants fairly across seasons?
	•	Would a binary “Service” vs. “Non-Service” label be better than my broader employment categories
	•	What statistical tests are best suited for comparing placement across groups with unequal sizes?

```{r setup}
library(tidyverse)
#install.packages('corrplot')
library(corrplot)
```

## Load the Data
```{r}
survivor_data <- read_csv("https://github.com/chelsea-collier/jour405_spring25/blob/c728c4d467755b6afb890fac20df0e32827823ba/project/Survivor_Data_402.csv")
head(md_turnout)
```
