---
title: "Multiple Regression Analysis: Maryland Voter Turnout"
name: "Chelsea Collier"
---

# Introduction

For my final project, I am analyzing a dataset that includes all contestants from Seasons 1 through 43 of the reality show Survivor. The dataset covers from Season 1 in 2000 to Season 43 in 2022, and includes demographic details, gameplay outcomes, and personal background information for each contestant.

The original dataset contained over 30 columns, but I’ve cleaned and narrowed it down to the best ones for my research questions. These include:
	•	profession: the contestant’s occupation, which I categorized into broader employment types such as Healthcare, Education, Business, etc.
	•	state: used to assign contestants to U.S. regions (East Coast, Midwest, South, West)
	•	region: derived from state data, this field is key to testing my hypothesis about East Coast performance
	•	finish and normalized_finish: these represent a contestant’s placement in their season, which is the main outcome variable
	•	num_season: the season number, which helps track changes across time
	•	birthdate, age, and gender: optional control variables for further analysis

#Research Questions

My main hypotheses are:
	1.	Contestants in service-related professions (e.g., education, healthcare, hospitality) place higher on average in Survivor
	2.	Contestants from the East Coast perform better in terms of final placement compared to those from other regions
	3.	Age Hypothesis: Contestants aged 25 to 35 are more likely to finish higher in the game compared to other age groups

#What’s Missing
	•	Income or education levels of contestants aren’t included, which might add context to occupation-based patterns
	•	A few state entries are missing or ambiguous, which affects accurate regional classification.
	•	I may also need to explore whether normalized vs. raw finish is the better metric for comparison across seasons with different numbers of contestants

#Questions
	•	Should I analyze raw finish (finish) or scaled placement (normalized_finish) to compare contestants fairly across seasons?
	•	Would a binary “Service” vs. “Non-Service” label be better than my broader employment categories
	•	What statistical tests are best suited for comparing placement across groups with unequal sizes?

```{r setup}
library(tidyverse)
#install.packages('corrplot')
library(corrplot)
```

## Load the Data
```{r}
survivor_data <- read_csv("https://github.com/chelsea-collier/jour405_spring25/blob/c728c4d467755b6afb890fac20df0e32827823ba/project/Survivor_Data_402.csv")
head(md_turnout)
```
# The Tests You Could Run (gpt used https://chatgpt.com/share/682a860e-8e18-800a-93eb-4f547d844e8d)
1. Comparing Mean Placement by Employment Category
	•	Test: One-way ANOVA or t-test (depending on number of groups)
	•	Why: I want to see whether contestants in service-related professions (e.g., healthcare, education, hospitality) place higher in Survivor compared to other occupational groups.
	•	Approach:
	•	I’ve already grouped contestants into employment categories.
	•	If I reduce this to “Service” vs. “Non-Service,” I will use a two-sample t-test.
	•	If I keep multiple categories (e.g., Education, Healthcare, Business, etc.), I’ll use a one-way ANOVA and follow up with Tukey’s HSD if significant.

2. Regional Differences in Placement
	•	Test: One-way ANOVA (East Coast, Midwest, South, West)
	•	Why: This will test if there’s a statistically significant difference in average placement across U.S. regions.
	•	Approach:
	•	Group contestants by region (already done).
	•	Compare normalized finish scores across these regions.

3. Does Age 25–35 Perform Best?
	•	Test: One-way ANOVA or post-hoc comparisons by age group
	•	Why: This will allow me to test whether contestants in the 25–35 age range perform better than those younger or older.
	•	Approach:
	•	Create age brackets (e.g., <25, 25–35, >35).
	•	Compare mean placements using ANOVA.
	•	Visualize with boxplots.
	
4. Regression Analysis
	•	Test: Linear regression
	•	Why: This can show how multiple variables (e.g., age, region, occupation) collectively predict performance in the game.
	•	Approach:
	•	Use normalized_finish as the dependent variable.
	•	Include predictors like age, gender, region, and occupation category.
	
	5. I would also likle to use a scatterplots to visualize the relationships of variables that may be correlated

